---
title: "Application Exercise"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: lab.css
link-citations: yes
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE     # for regression output
)
```


1. Open Rstudio Server Pro
2. Create a new project
3. Click File > New File > R Markdown
4. Delete everything except for 
```
---
title: "Untitled"
output: html_document
---
```
5. Give your work a sensible title
6. After adding your derivation to the file (remember you can incorporate $\LaTeX$ by wrapping in `$$`), knit it.
6. When you've finished, upload the .html file to Canvas


* Fit a linear model using the `mtcars` data frame predicting miles per gallon (`mpg`) from (`wt`)
* Pull out the coefficients and confidence intervals using the `tidy()` function demonstrated. How do you interpret these?
*  Using the linear model you fit previously (`mpg` from `wt` using the `mtcars` data) - calculate the p-value for the coefficient for weight? Interpret this value. What is the null hypothesis? What is the alternative hypothesis? Do you reject the null?
* Using the model previously fit (using the `mtcars` data predicting miles per gallon from weight), pull out the F-statistic and $R^2$ using the `glance()` function. Interpret these values.

Below is some code to get you started:

```{r, eval = FALSE}
library(tidymodels)
library(tidyverse)
```

