---
title: "Application Exercise"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: lab.css
link-citations: yes
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE     # for regression output
)
```


1. Open Rstudio Server Pro
2. Create a new project
3. Click File > New File > R Markdown
4. Delete everything except for 
```
---
title: "Untitled"
output: html_document
---
```
5. Give your work a sensible title
6. When you've finished, upload the .html file to Canvas


5000 women were enrolled in a study and were randomly assigned to receive either letrozole or a placebo. The primary response variable of interest was disease-free survival.

   |letrozole	| placebo	| total	
---|----------|---------|------
death or disease|	185	|341|	526
no death or disease|	2390|	2241|	4631
total|	2575|	2582|	5157

* calculate the odds ratio of death or disease in the placebo group versus the treatment group
* calculate the odds ratio of no death or disease in the treatment group versus the placebo group
* calculate the odds ratio of death or disease in the treatment group versus the placebo group


* Using the `Default` data from the `ISLR` package, fit a logistic regression model predicting whether a customer `defaults` with whether they are a `student` and their current `balance`.

Here is some code to get you started:

```{r}
library(ISLR)
data("Default")
```

