---
title: "Application Exercise"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: lab.css
link-citations: yes
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE     # for regression output
)
```


1. Open Rstudio Server Pro
2. Create a new project
3. Click File > New File > R Markdown
4. Delete everything except for 
```
---
title: "Untitled"
output: html_document
---
```
5. Give your work a sensible title
6. When you've finished, upload the .html file to Canvas

* In R, find a design matrix `X` where it is not possible to calculate $\hat\beta$ 

```{r, eval = FALSE}
solve(t(X) %*% X) %*% t(X) %*% y
```

* Calculate the determinant of the following matrices in R using the `det()` function:

$$\mathbf{A} = \begin{bmatrix} 1 & 2 \\ 4 & 5 \end{bmatrix}$$

$$\mathbf{B} = \begin{bmatrix} 1 & 2 & 3 \\ 3 & 6 & 9 \\ 2 & 5 & 7\end{bmatrix}$$
* Are these both invertible?

* Find $\hat\beta$ that minimizes this:

$$(\mathbf{y}-\mathbf{X}\beta)^T(\mathbf{y}-\mathbf{X}\beta)+\lambda\beta^T\beta$$