---
title: "Lab 03 - Logistic Regression"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: lab.css
link-citations: yes
---

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(ISLR)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE     # for regression output
)
```

**Due:** Tuesday 2020-10-06 at 5pm

# Getting started

- Go to today's module in Canvas - click "Download Lab .Rmd starter file"
- Go to RStudio Pro. Create a **New Project** by clicking File > New Project > New Directory. Name the directory something sensible (for example `03-lab-logistic`)
- Open a new .Rmd file and you're ready to go!


# Packages

In this lab we will work with four packages: `ISLR` which is a package that accompanies your textbook, `tidyverse` which is a collection of packages for doing data analysis in a "tidy" way, `tidymodels` a collection of packages for statistical modeling, and `GGally`, a package to help us visualize the data.

```{r eval = FALSE}
library(tidyverse) 
library(tidymodels)
library(ISLR)
library(GGally)
```

# Data

For this lab, we are using the `Smarket` data from the `ISLR` package.

# Exercises

## Conceptual questions

1. Using algebra, rearrange equation (1) below to get equation (2). Show all steps. _Be sure to use Latex, remember you can insert a Latex equation by wrapping it in `$$`._

Equation (1)

$$p(X)=\frac{e^{\beta_0+\beta_1X}}{1+e^{\beta_0+\beta_1X}}$$

Equation (2)

$$\textrm{log}\left(\frac{p(X)}{1-p(X)}\right)=\beta_0+\beta_1X$$

2. Suppose that an individual has a 23% chance of defaulting on their credit card payment. What are the odds that they will default?

## Logistic Regression

3. For this lab we are using the `Smarket` data. Examine this data set - how many observations are there? How many columns? What are the variables?

4. Let's look at the correlation between all of the variables. To do this, if you haven't done so already, we need to install the `GGally` package. Run the following code in your Console **one time**. What can you learn from this visualization? Which pair of variables have the highest correlation?

```{r, eval = FALSE}
install.packages("GGally")
```

```{r, eval = FALSE}
ggpairs(Smarket, lower = list(combo = wrap(ggally_facethist, binwidth = 0.5)), progress = FALSE)
```
5. **Inference** Fit a logistic regression model to predict `Direction` using `Lag1`, `Lag2`, `Lag3`, `Lag4`, `Lag5`, and `Volume`. Show a table that contains the coefficients and p-values along with the confidence intervals for each of the 6 predictors. Which predictor has the smallest p-value? Interpret the coefficient, confidence interval, and p-value for this predictor.

6. **Inference** Exponentiate the results from Exercise 5. Interpret the _odds ratio_ for the same predictor you selected in Exercise 5.

7. **Prediction** Using 5-fold cross validation, fit the same logistic regression model as Exercise 5. What is the test Accuracy for this model? Interpret this result.


8. **Inference** Fit a logistic regression model to predict `Direction` using only `Lag1` and `Lag2`. Show a table that contains the coefficients and p-values along with the confidence intervals for each of the 2 predictors. Which predictor has the smallest p-value? Interpret the coefficient, confidence interval, and p-value for this predictor.

9. **Inference** Exponentiate the results from Exercise 8. Interpret the _odds ratio_ for the same predictor you selected in Exercise 8.

8. **Prediction** Using 5-fold cross validation, fit the same logistic regression model as Exercise 8. What is the test Accuracy for this model? Interpret this result.


9. If you had to choose between the model fit in Exercise 5 and the one fit in Exercise 8, which would you choose? Why?
