---
title: "Lab 05 - Nonlinear models"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: lab.css
link-citations: yes
---

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(ISLR)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE     # for regression output
)
```

**Due:** Tuesday 2020-11-03 at 5:00pm


# Packages

In this lab we will work with three packages: `ISLR` for the data, `tidyverse` which is a collection of packages for doing data analysis in a "tidy" way and `tidymodels` for statistial modeling.

If you'd like to run your code in the Console as well you'll also need to load the packages there. To do so, run the following in the console. 

```{r eval = FALSE}
library(tidyverse) 
library(tidymodels)
library(ISLR)
```

Note that the packages are also loaded with the same commands in your R Markdown document.

# Data

For this lab, we are using `Wage` data from the `ISLR` package.

# Exercises

1. Examine the `Wage` data set from the `ISLR` package. What are the variables? How many observations are there?

2. Create a linear model specification, setting the engine to `lm`. Call this model specification `linear_spec`.

3. Create a recipe using the `Wage` data from the `ISLR` package. We want to predict the variable `wage` from `age`, `health_ins`, `jobclass`, `education`, and `race`. We want age to be fit using a natural spline. Use `tune()` to decide how many degrees of freedom to use for the `age` variable. (_In this exercise, you are just creating the recipe_).

4. Use `tune_grid()` to fit the linear model specified in Exercise 2 with the recipe created in Exercise 3 using 10-fold cross validation, similar to Lab 04. Choose the model with the lowest RMSE. How many degrees of freedom were used for the age natural spline for this best model? Report the RMSE for this model as well as the chosen degrees of freedom.

5. Create a recipe using the `Wage` data from the `ISLR` package. We want to predict the variable `wage` from `age`, `health_ins`, `jobclass`, `education`, and `race`. We want age to be fit using a polynomial. Use `tune()` to decide how many degrees the polynomial should have for the `age` variable. (_In this exercise, you are just creating the recipe_).

6. Use `tune_grid()` to fit the linear model specified in Exercise 5 with the recipe created in Exercise 3 using 10-fold cross validation, similar to Lab 04. Choose the model with the lowest RMSE. What degree polynomial was used for age for this best model? Report the RMSE for this model as well as the chosen degree.

7. If the goal is prediction, which model would you prefer, the one fit in Exercise 4 or the one fit in Exercise 6? Using your chosen model, examine whether ridge, lasso, or elastic net would provide a better fit. Describe your results.





